---
title: "p8105_hw1_ng3005"
output: github_document
date: "September 20, 2025"
---

## Problem 1

```{r}
library(tidyverse)
library(knitr)
library(moderndive)
data("early_january_weather")
```

To write a short description of the dataset, we can visualize the first few rows

```{r}
head(early_january_weather)
```

This dataset contains 15 variables: 'origin', 'year', 'month', 'day', 'hour', 'temp', 'dewp', 'humid', 'wind_dir', 'wind_speed', 'wind_gust', 'precip', 'pressure', 'visib', and 'time_hour'

```{r}
nrow(early_january_weather)
ncol(early_january_weather)
```

Using `nrow` and `ncol`, there are 358 rows and 15 columns in the dataset.

```{r}
mean(early_january_weather$temp)
```

Using the `mean()` function, the mean temperature in the dataset is 39.58212.

#### Let's create a scatterplot of `temp` (y) vs `time_hour` (x)

```{r}
ggplot(
  early_january_weather, 
  aes(x = time_hour, y = temp, color = humid)) + geom_point()
```

Overall, this scatterplot shows an upward trend in temperature as time progresses. The points also seem to oscillate up and down in temperature likely fluctuating due to daytime and nighttime. There is also an apparent increase in humidity around Jan 12 to Jan 14. 

#### Let's export the above scatterplot to the project directory

```{r}
ggsave("p8105_hw1_ng3005_scatter.pdf", height = 4, width = 6)
```

## Problem 2

#### Create dataframe

```{r}
hw1_df = tibble(
  random_sample = rnorm(10),
  vec_positive = random_sample > 0,
  vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  vec_factor = factor(
      rep(c("level_1", "level_2", "level_3"), length.out = 10),
      levels = c("level_1", "level_2", "level_3")
  )
)

hw1_df
```


