---
title: "p8105_hw1_ng3005"
author: "Nelson Gaillard"
UNI: "ng3005"
output: github_document
date: "September 20, 2025"
---

## Problem 1

```{r}
library(tidyverse)
library(knitr)
library(moderndive)
```

Load the dataset from the the moderndive library

```{r}
data("early_january_weather")
```

To write a short description of the dataset, we can visualize the first few rows

```{r}
head(early_january_weather)
```

This dataset contains 15 variables: 'origin', 'year', 'month', 'day', 'hour', 'temp', 'dewp', 'humid', 'wind_dir', 'wind_speed', 'wind_gust', 'precip', 'pressure', 'visib', and 'time_hour'. It is important to note that the variable `origin` identifies the weather station from which each row's data was collected. Both `temp` and `dewp` (temperature and dewpoint) are in units of degrees Fahrenheit. `wind_dir` is in units of degrees. `wind_speed`, and `wind_gust` are in units of miles per hour (mph). `precip`, `pressure`, and `visib` are in units of inches, millibars, and miles, respectively. `time_hour` identifies the date and hour of the recording.

Dataset `early_january_weather` contains `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns. The mean temperature in the dataset is `r mean(pull(early_january_weather, temp))` degrees Fahrenheit. 

```{r}
nrow(early_january_weather)
ncol(early_january_weather)
```

Using `nrow` and `ncol`, there are 358 rows and 15 columns in the dataset.

```{r}
mean(early_january_weather$temp)
```

Using the `mean()` function, the mean temperature in the dataset is 39.58212.

#### Let's create a scatterplot of `temp` (y) vs `time_hour` (x)

```{r}
ggplot(
  early_january_weather, 
  aes(x = time_hour, y = temp, color = humid)) + geom_point()
```

Overall, this scatterplot shows an upward trend in temperature as time progresses. The points also seem to oscillate up and down in temperature likely fluctuating due to daytime and nighttime. There is also an apparent increase in humidity around Jan 12 to Jan 14. 

#### Let's export the above scatterplot to the project directory

```{r}
ggsave("p8105_hw1_ng3005_scatter.pdf", height = 4, width = 6)
```

## Problem 2

#### Create dataframe

```{r}
hw1_df = tibble(
  random_sample = rnorm(10),
  vec_positive = random_sample > 0,
  vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  vec_factor = factor(
      rep(c("level_1", "level_2", "level_3"), length.out = 10),
      levels = c("level_1", "level_2", "level_3")
  )
)

hw1_df
```

Let's try to take the mean of each variable

```{r}
mean(hw1_df $ random_sample)

mean(hw1_df $ vec_positive)

mean(hw1_df $ vec_char)

mean(hw1_df $ vec_factor)
```

